---
title: "Marketing Hw1"
output: pdf_document
date: "2024-02-23"
---

# Before starting the homework itself let's install the necesary packages

```{r cars}
libs<-c('ggplot2','ggpubr','knitr','diffusion')
load_libraries<-function(libs){
  new_libs <- libs[!(libs %in% installed.packages()[,"Package"])] 
  if(length(new_libs)>0) {install.packages(new_libs)}
  lapply(libs, library, character.only = TRUE)
}
load_libraries(libs)
```

## Every year, Time magazine publishes a list of the best 100 innovations of that year.  Go to the list, choose an innovation, and put the link of the selected product. I selected a product, which is a "A Longer-lasting Laptop" Framework Laptop 16. It is a new laptop that can be continually modified by swapping out its parts, which will let users upgrade the CPU and graphics—and keep older machines out of landfills. Thinking about look-alike innovation from the past I decided to take simple laptops which at that time where the boom in the market and changed a lot of things in humans' lives. For this reason, I found a data related to revenue of the laptops industry worldwide from 2019-2028, but I will use just the subset of it and make predictions based on that. I guess the innovation of the laptop was world-changing and year after year even more interesting and inovative sort of it is produced and introduced to the market. 


```{r}
library(readxl)
excel_file <- "path/to/your/excel/file.xlsx"

# Read the Excel file
data <- read_excel(excel_file, col_names = FALSE)

# Print the data
print(data)

```

